---
title: "NVTA_Results"
author: "Volpe Center (Sarah Lohmar, Dan Flynn, Eric Englin)"
date: "6/24/2021"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Metropolitan Data Visualization 

The Volpe team started by looking at the daily VMT for all scenarios. This is done by combining seven DVMT outputs by location and type of mode. Percent change is calculated relative to the base scenario value.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)

# This sections sets the location of the VisionEval installation for each of the Volpe team members.
# The result is that the object `ve.runtime` is the full path to where VisionEval is installed, and the results from the NVTA scenarios are located. 
ve.runtime <- ifelse(grepl('Flynn', normalizePath('~/')), 
                     file.path(dirname('~/'), 'Desktop/VE_4-1-0/'),
                      ifelse(grepl('lohmar', normalizePath('~/')), 
                       'C:/VisionEval',
                         ifelse(grepl('englin', normalizePath('~/')), 
                         file.path('C:/Users/eric.englin/Desktop/VisionEval/4_04_v3/'),
                          NA)))

# Make a directory to save JPEG outputs as well
fig_save = file.path(ve.runtime, 'models', 'NVTA_Figures')
if(!dir.exists(fig_save)) { dir.create(fig_save) }

knitr::opts_knit$set(root.dir = ve.runtime)

library(tidyverse)
library(plotly) #install.packages('plotly')

```


## DVMT Plots

### Percent change 

```{r dvmt_rel}
readfile <- read.csv(file.path(ve.runtime, 'models', 'Scenario_Metrics_Marea_with_Base.csv'))

#info for regional dvmt graphic 
dvmt <- readfile %>%
  select(modelName, contains('Dvmt')) %>%
  rowwise() %>%
  mutate(modelName = sub('VERSPM_', '', modelName)) %>%
  mutate(regionalDVMT = sum(c_across(contains('Dvmt'))))

dvmt$modelName[dvmt$modelName == 'NVTA'] = "Base"

dvmtbase = as.numeric(dvmt %>% filter(modelName == 'Base') %>% select(regionalDVMT))


dvmt <- dvmt %>%
  mutate(regDVMTAbsDif = regionalDVMT - dvmtbase,
         regDVMTPerChange = (regDVMTAbsDif/dvmtbase)*100,
         color = ifelse(regDVMTPerChange > 0, "pos","neg"),
         pct_text = paste('Scenario:', modelName, '\n', round(regDVMTPerChange, 2), '%'),
         mi_text = paste('Scenario:', modelName, '\n', format(round(regDVMTAbsDif, 2),
                                                              scientific = F, 
                                                              big.mark = ','), 'mi/day'))

gp1 <- ggplot(dvmt, aes(x = modelName, y = regDVMTPerChange, fill = color,
                        text = pct_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Percent Change Regional DVMT", x = "Scenario", y = "Percent Change") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Save the output figure to disk for inclusion in a manuscript
ggsave(file.path(fig_save, 'DVMT_Rel_Change.jpeg'))

# And plot dynamically for viewing.
ggplotly(gp1, tooltip = 'text')
```

### Absolute change

```{r dvmt_abs}
gp2 <- ggplot(dvmt, aes(x = modelName, y = regDVMTAbsDif, fill = color,
                        text = mi_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Absolute Change Regional DVMT (mi/day)", x = "Scenario", y = "Absolute Change")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'DVMT_Abs_Change.jpeg'))

ggplotly(gp2, tooltip = 'text')
```

## Energy Usage plots

### Truck energy usage relative

```{r truck_energy_rel}

#info for commercial heavy truck gge (sum columns k-m)
# use `matches()` to use regular expressions first to find variable names that start with 
# ComSvc or HvyTrk, and then a second one to find only those which also have GGE at the end
gge <- readfile %>%
  select(modelName, matches('^(ComSvc|HvyTrk)')) %>%
  select(modelName, matches('GGE$')) %>%
  rowwise() %>%
  mutate(modelName = sub('VERSPM_', '', modelName)) %>%
  mutate(ComHvyTruckGGE = sum(c_across(contains('GGE'))))

gge$modelName[gge$modelName == 'NVTA'] = "Base"

ggebase = as.numeric(gge %>% filter(modelName == 'Base') %>% select(ComHvyTruckGGE))

gge <- gge %>%
  mutate(ComHvyTruckGGEAbsDif = ComHvyTruckGGE - ggebase,
         ComHvyTruckGGEPerChange = (ComHvyTruckGGEAbsDif/ggebase)*100,
         sign = ifelse(ComHvyTruckGGEPerChange > 0, "pos","neg"),
         pct_text = paste('Scenario:', modelName, '\n', round(ComHvyTruckGGEPerChange, 2), '%'),
         mi_text = paste('Scenario:', modelName, '\n', format(round(ComHvyTruckGGEAbsDif, 2),
                                                              scientific = F, 
                                                              big.mark = ','), 'Gasoline gallone equiv.'))

# View(gge)

gp3 <- ggplot(gge, aes(x = modelName, y = ComHvyTruckGGEPerChange, fill = sign,
                       text = pct_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Percent Change Commercial Service and  Heavy Truck GGE", x = "Scenario", y = "Percent Change")+
  theme(axis.text.x = element_text(angle = 45))

ggsave(file.path(fig_save, 'Truck_GGE_Pct_Change.jpeg'))

ggplotly(gp3, tooltip = 'text')
```


# Household Data Visualization

```{r household_prep, cache = T}
# Use `cache = T` for faster loading
hhfile <- readr::read_csv(file.path(ve.runtime, 'models', 'Scenario_Metrics_Hh_with_Base.csv'))

baseNVTA <- hhfile %>%
  filter(modelName == "VERSPM_NVTA") 

scenario_values <- hhfile %>%
  filter(modelName != 'VERSPM_NVTA') %>% # exclude the base scenario
  mutate(modelName = sub('VERSPM_', '', modelName)) %>%
  group_by(modelName) %>%
  summarize(sum_Hh_GGE = sum(DailyGGE),
              sum_Hh_kWh = sum(DailyKWH),
              sum_Hh_CO2e = sum(DailyCO2e))
  
base_values <- hhfile %>%
  filter(modelName == "VERSPM_NVTA") %>%
  mutate(modelName = 'NVTA_Base') %>%
  summarize(base_Hh_GGE = sum(DailyGGE),
            base_Hh_kWh = sum(DailyKWH),
            base_Hh_CO2e = sum(DailyCO2e))

# We use base_values$... to compare the scenario values with the corresponding base values data frame

scenario_values <- scenario_values %>%
  mutate(GGEPerChange = ((sum_Hh_GGE - base_values$base_Hh_GGE) / base_values$base_Hh_GGE) * 100,
         KWHPerChange = ((sum_Hh_kWh - base_values$base_Hh_kWh) / base_values$base_Hh_kWh) * 100,
         CO2PerChange = ((sum_Hh_CO2e - base_values$base_Hh_CO2e) / base_values$base_Hh_CO2e) * 100,
         gge_color= ifelse(GGEPerChange > 0, "pos","neg"),
         gge_pct_text = paste('Scenario:', modelName, '\n', round(GGEPerChange, 2), '%'),
         gge_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_GGE, 2),
                                                              scientific = F, 
                                                              big.mark = ','), 'Gasoline gallone equiv.'),
         kwh_color= ifelse(KWHPerChange > 0, "pos","neg"),
         kwh_pct_text = paste('Scenario:', modelName, '\n', round(KWHPerChange, 2), '%'),
         kwh_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_kWh, 2),
                                                              scientific = F, 
                                                              big.mark = ','), 'kWh'),
         co2_color= ifelse(CO2PerChange > 0, "pos","neg"),
         co2_pct_text = paste('Scenario:', modelName, '\n', round(CO2PerChange, 2), '%'),
         co2_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_CO2e, 2),
                                                              scientific = F, 
                                                              big.mark = ','), 'g CO[2] equivalent'))
         

gp4 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_GGE,
                                   text = gge_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional daily gasoline gallon equivalents", x = "Scenario", y = "GGE/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_GGE_Total.jpeg'))

ggplotly(gp4, tooltip = 'text')

gp5 <- ggplot(scenario_values, aes(x = modelName, y = GGEPerChange,
                                   text = gge_text, fill = gge_color)) + 
  geom_bar(stat="identity") + theme_minimal() +
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Regional daily gasoline gallon equivalents % Change", x = "Scenario", y = "GGE/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_GGE_Percent_Change.jpeg'))

ggplotly(gp5, tooltip = 'text')

```
### Electricity usage

```{r household_kwh}
gp6 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_kWh,
                            text = kwh_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional KWH", x = "Scenario", y = "kWh/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_kWh_Total.jpeg'))


ggplotly(gp6, tooltip = 'text')

gp7 <- ggplot(scenario_values, aes(x = modelName, y = KWHPerChange,
                            text = kwh_text, fill = kwh_color)) + 
  geom_bar(stat="identity") + theme_minimal() +
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Regional KWH Percent Change", x = "Scenario", y = "kWh/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



ggsave(file.path(fig_save, 'Regional_kWh_Percent_Change.jpeg'))

ggplotly(gp7, tooltip = 'text')
```

### Emissions

```{r household_emissions}
gp8 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_CO2e,
                            text = co2_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional CO[2]e", x = "Model Name", y ="g CO[2] equivalent / Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggsave(file.path(fig_save, 'Regional_CO2e_Total.jpeg'))

ggplotly(gp8, tooltip = 'text')

gp9 <- ggplot(scenario_values, aes(x = modelName, y = CO2PerChange,
                            text = co2_text, fill = co2_color)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional CO[2]e Percent Change", x = "Model Name", y ="g CO[2] equivalent / Day")+
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggsave(file.path(fig_save, 'Regional_CO2e_Percent_Change.jpeg'))

ggplotly(gp9, tooltip = 'text')


```

# Validation 

The Volpe team also realized the need to validate the VisionEval output. 

#### MWCOG DVMT

The VDOT team sent along daily VMT values in 2021 and 2045 from the MWCOG MOE Base model. These results include household car VMT, business car VMT, and car travel from other locations (such as motorists who live outside the area and are passing through it.

Here are the raw results from the MWCOG base model in 2045:

       Jurisdiction  Freeway     Expressway  Major Art.  Minor Art.   Collector     Ramp       All
-------------------- ----------- ----------- ----------- ----------- ----------- ----------- -----------
 Fairfax County      12,517,806   5,910,760   6,219,736   1,419,677   1,256,563     488,882  27,813,424
 City of Fairfax              0     284,867      99,406       9,144           0           0     393,417
 Falls Church                 0     106,172      40,693       4,605           0           0     151,470
 
 
CROSSTAB ROW=FLAG COL=FTYPE VAR=_24VMTTRK
                                                                                                       
       Jurisdiction  Freeway     Expressway  Major Art.  Minor Art.   Collector     Ramp       All
-------------------- ----------- ----------- ----------- ----------- ----------- ----------- -----------
Fairfax County        3,071,425     988,206     919,784     212,966     331,425      90,603   5,614,408
City of Fairfax               0      37,640      12,957         774           0           0      51,371
Falls Church                  0      18,944       5,575         425           0           0      24,944


When you add all of the numbers up, you get 34,049,034 daily VMTs. This is about 4,000,000 higher than the VisionEval dvmt values. The chart below shows the percent differences with each of our scenarios. 

```{r dvmt_plot_validation}

ffx_county_car_2045 = 27813424
ffx_city_car_2045 = 393417
fall_church_car_2045 = 151470
ffx_county_truck_2045 = 5614408
ffx_city_truck_2045 = 51371
fall_church_truck_2045 = 24944

MWCOG_model_dvmt = ffx_county_car_2045+ffx_city_car_2045+fall_church_car_2045+ffx_county_truck_2045+ffx_city_truck_2045+fall_church_truck_2045

#trip_generation_summary_motorized_dvmt = 3649551
#trip_generation_summary_non_motorized_dvmt = 317247

dvmt$regDVMTAbsDif = dvmt$regionalDVMT - MWCOG_model_dvmt
dvmt$regDVMTPerChange = (dvmt$regDVMTAbsDif/MWCOG_model_dvmt)*100
dvmt$color <- ifelse(dvmt$regDVMTPerChange < 0, "red","green")

dvmtplot1 <- data.frame(Model = dvmt$modelName,PercentChange = dvmt$regDVMTPerChange, 
                        color = dvmt$color)

ggplot(dvmtplot1, aes(x=Model, y=PercentChange, fill = color)) + 
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Percent Change VisionEval DVMT and MWCOG MOES Base Model for 2045", x = "Model Name", y = "Percent Change")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


Here is a table showing some of the validation steps taken:

| File from the Travel Demand Model | Value       | Output from NOVA model | Output from VisionEval (Base Model) | Percent Difference |
| --------------------------------- | ----------  | ---------------------- | ----------------------------------  | ------------------ |
| 2019i4_Trip_Generation_Summary.txt | HH DVMT    | 34,049,034             | 30,040,347                          | 5-15%              |
| 2019i4_Trip_Generation_Summary.txt | Truck DVMT | 5,690,723              | 1,206,452                           | 70-80%              |

